# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
- Extract zip activity file from forked repository to generate the activity.csv file.
Read activity.csv into a dataframe.
```{r }
data <- read.csv("I:/Coursera/Reproducible research/git/RepData_PeerAssessment1/activity/activity.csv")

```
- Look at the data  
```{r }
class(data)
head(data)
str(data)
```
## What is mean total number of steps taken per day?
- create histogram of total steps per day  
```{r histogram, fig.height=4}
sums <- tapply(data$steps, data$date, sum)
hist(sums)
```
---
- calculate mean and median of steps per day, removing na  
```{r }
print(mean(sums, na.rm = T))
print(median(sums, na.rm = T))
```
## What is the average daily activity pattern?
- time series plot of 5 min intervals and average steps taken  
```{r plot}
step_av<-tapply(data$steps, data$interval, mean, na.rm = T)
plot(unique(data$interval), step_av,  type = "l", xlab = "interval", ylab = "step average")
```
---
- determine which interval has maximum number of steps  
```{r }
attr(which.max(step_av), "names")
```
## Imputing missing values
- total missing values
```{r }
na_steps <- is.na(data$steps)
missing <- data[na_steps,]
length(missing$steps)
```
- fill missing values and create new dataset 
use means of intervals over all days 
```{r }
for (i in 1:length(data[,1])) {
    if (is.na(data[i,1])) {
        est_step <- step_av[name=as.character(data[i,3])]
        names(est_step) <- NULL
        data[i,1]<-est_step
    }
}
print (head(data, 20))
```
- plot histogram of steps per day  
and calculate mean and median steps per day
```{r }
sums1 <- tapply(data$steps, data$date, sum)
head(sums1)
hist(sums1)
print(mean(sums1))
print(median(sums1))
```
- compare mean and median of original and revised (replaced NA) datasets
```{r }
library(data.table)
dt <- data.table(orig = c(mean(sums, na.rm=T), median(sums, na.rm=T)), rev = c(mean(sums1), median(sums1)))
print(dt)
```
## Are there differences in activity patterns between weekdays and weekends?
- add column to dataset for weekdays based on date column
```{r }
days<- factor(weekdays(as.Date(data$date)))
levels(days)[levels(days) == "Monday"] <-"weekday"
levels(days)[levels(days) == "Tuesday"] <-"weekday"
levels(days)[levels(days) == "Wednesday"] <-"weekday"
levels(days)[levels(days) == "Thursday"] <-"weekday"
levels(days)[levels(days) == "Friday"] <-"weekday"
levels(days)[levels(days) == "Saturday"] <- "weekend"
levels(days)[levels(days) == "Sunday"] <- "weekend"

#head(days)
data1<-cbind(data, days)
library(lattice)
```
- calculate mean of steps for each interval using aggregate, assign column names, and plot
```{r }
ag_data <- aggregate(data1$steps ~ data1$interval + data1$days, list(data1$days), mean)  

colnames(ag_data)<-c("interval", "days", "steps")

xyplot(steps ~ interval | days, data = ag_data, layout = c(1,2), type = "l")

```
